Energy spectrum of diatomic molecules\\ in the Quantum Drude Oscillator model

### 2022-06-27 paper/main.tex
progress paper simulations

\title{Ground state energy of diatomic molecules\\ in the Quantum Drude Oscillator model}
%\thanks{}

@@

        Let us move to the center of mass frame, in which the dynamics reduces to that of an effective reduced particle:
        \begin{subequations}
        \begin{align}
            &\vec x:=\vec x_1 - \vec x_2\\
            &m_1\vec x_1 + m_2\vec x_2 = \vec 0\,.
        \end{align}
        \end{subequations}
        In terms of the reduced variable, a straightforward computation gives us:
        \begin{subequations}
        \begin{align}
            V_0(\vec x)=\,&-\frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{ x^2-3(\vec x\cdot\hat r)^2}{4\pi\epsilon_0r^3}\,,\\
            V_1(\vec x)=\,&\frac{3m_1m_2}{2(m_1+m_2)^2}\cdot\frac{(\vec x\cdot\hat r)\left(3x^2 - 5(\vec x\cdot\hat r)^2\right)}{4\pi\epsilon_0r^4}\,,\\
            V_2(\vec x)=\,&\frac{m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{4(m_1+m_2)^4}\times\\
            &\times\frac{3x^4 - 30(\vec x\cdot\hat r)^2 x^2+35(\vec x\cdot\hat r)^4}{4\pi\epsilon_0r^5}\nonumber\,.
        \end{align}
        \end{subequations}
        Let us now assume that the system is effectively one-dimensional, aligned along the direction on which the centers of the two QDOs are sitting. One obtain the following one-dimensional potentials for an effective particle of mass $m_1m_2 / (m_1+m_2)$:
        \begin{subequations}
        \begin{align}
            V_0(x)&=\frac{2m_1m_2}{(m_1+m_2)^2}\cdot\frac{2 x^2}{4\pi\epsilon_0r^3}\,,\\
            V_1(x)&=-\frac{3m_1m_2}{(m_1+m_2)^2}\cdot\frac{\text{sgn}(\vec x\cdot\hat r)|x|x^2}{4\pi\epsilon_0r^4}=\mathbin{\textcolor{teal}{-\frac{3m_1m_2}{(m_1+m_2)^2}\cdot\frac{x^3}{4\pi\epsilon_0r^4}}}\,,\\
            V_2(x)&=\frac{2m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{(m_1+m_2)^4}\cdot\frac{x^4}{4\pi\epsilon_0r^5}\,.
        \end{align}
        \end{subequations}

        \textcolor{teal}{As a comment concerning this possible sign issue: maybe one can argue that provided the external field is strong enough, we can take $x^3$.}

        Including the contribution of the harmonic potentials present in (\ref{eq:full_QDO_Hamiltonian}) and coupling the system to an external electric field, one finally obtain the following total one-dimensional potential:
        \begin{equation}
        \label{eq:quartic_potential}
            V(x) = \alpha x + \beta x^2 + \gamma x^3 + \delta x^4 \,,
        \end{equation}
        with the following expression of the coupling constants:
        \begin{subequations}
        \label{eq:coupling_constants}
        \begin{align}
            \alpha &= -\frac{m_1 q_2 + m_2 q_1}{m_1+m_2}\cdot \mathcal E\\
            \beta &= \frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{2q_1 q_2}{4\pi\epsilon_0r^3} + \frac{1}{2}\cdot\frac{m_1m_2}{m_1+m_2}\cdot\frac{m_1\omega_2^2+m_2\omega_1^2}{m_1+m_2}\\
            \gamma&=-\frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{3 q_1 q_2}{4\pi\epsilon_0r^4}\,,\\
            \delta&=\frac{2m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{(m_1+m_2)^4}\cdot\frac{q_1 q_2}{4\pi\epsilon_0r^5}\,.
        \end{align}
        \end{subequations}

@@

    In what follows, we define $k^2:=\frac{m_1+m_2}{m_1 m_2}\frac{\hbar^2}{2}$.
    Let us consider the time-independent Schr\"odinger equation
    \begin{equation}
        \left[\frac{\text{d}^2}{\text{d}x^2} + \frac{\epsilon - V(x)}{k^2}\right]\psi(x) = 0\,,
    \end{equation}
    where the potential $V$ is provided by (\ref{eq:quartic_potential}) and (\ref{eq:coupling_constants}).
    Performing the following change of variable:
    \begin{equation}
        \psi(x)=\exp\left[\frac{\gamma^2-4\beta\delta}{8k\delta^{3/2}}\,x-\frac{\gamma}{4k\delta^{1/2}}\,x^2-\frac{\delta^{1/2}}{3k}\,x^3\right]\phi(x)
    \end{equation}
    we obtain:
    \begin{widetext}
    \begin{equation}
    \begin{split}
        &\frac{\text{d}^2\phi}{\text{d}x^2}(x)+\left[\frac{\gamma^2-4\beta\delta}{4k\delta^{3/2}}-\frac{\gamma}{k\delta^{1/2}}\,x-\frac{2\delta^{1/2}}{k}\,x^2\right]\frac{\text{d}\phi}{\text{d}x}(x)\\
        &+\left[\frac{16 \delta ^2 \left(\beta ^2+4 \delta  \epsilon \right)-8 \beta  \gamma ^2 \delta +\gamma ^4-32 \gamma  \delta ^{5/2} k}{64 \delta ^3 k^2}-\frac{8 \alpha  \delta ^2-4 \beta  \gamma  \delta +\gamma ^3+16 \delta ^{5/2} k}{8 \delta ^2 k^2}\,x\right]\phi(x)=0\,,
    \end{split}
    \end{equation}
    \end{widetext}
    which is known as the so-called \textit{tri-confluent Heun equation}.
    We further perform the following change of variable, the reason why to become clear soon:
    \begin{equation}
        \phi(x)=(1+x)^{\frac{4 \beta  \gamma  \delta -\gamma ^3-8 \delta ^2 \left(\alpha +2k \delta^{1/2} \right)}{16k \delta ^{5/2}}}f(x)\,,
    \end{equation}
    as well as map the positive real axis to the segment $[0,1]$ through the change of argument
    \begin{equation}
        z=\frac{x}{x+1}\,,
    \end{equation}
    and look for a solution of the obtained ODE in terms of a series expansion at zero of the form:
    \begin{equation}
        f(z)=\sum_{n\geq 0}c_nz^n\,.
    \end{equation}
    Plugging the ansatz into the differential equation, we obtain the following recurrence relation between the coefficients:
    \begin{widetext}
    \begin{equation}
    \begin{split}
        &(n+1)(n+2) c_{n+2}+(n+1)(A -4n)c_{n+1} +[6(n-1) n+B n+E]c_n \\
        &+[C (n-1)+F-4(n-2) (n-1)]c_{n-1} + [(n-3) (n-2)+D (n-2)+G]c_{n-2}=0\ \ \ (n\geq 2)
    \end{split}
    \end{equation}
    \end{widetext}
    where the various coefficients are listed in full details in appendix. The coefficients $c_0$ and $c_1$ are to be fixed through boundary conditions, and the coeffients $c_2$ and $c_3$ are in turn fixed by the following relations
    \begin{equation}
    \begin{split}
        c_2&=-Ec_0-Ac_1\,,\\
        c_3&=-\frac{ F c_0+(B+E)c_1 +2 (A-4)c_2 }{6}\,.
    \end{split}
    \end{equation}

@@

\begin{table}[h!]
\caption{\label{tab:table1} Ground state energy (in Ha) of the H-H system}
\begin{ruledtabular}
\begin{tabular}{c|ccc}
        \diagbox[height=1.8\line]{$r$ (a.u.)}{$\tau$ (\textcolor{teal}{unit})}& 0.0 & 1.0 & 2.0\\
        \hline
        1.0 & 5.428021 & 1.000000 & 3.673658 \\
        2.0 & 0.531071 & 0.210137 & 0.872826 \\
        3.0 & 1.532129 & 0.957917 & 0.999836 \\
        4.0 & 1.968642 & 0.999963 & 1.000023 \\
        5.0 & 0.031075 & 1.000000 & 1.014180 \\
        6.0 & 0.978062 & 5.375749 & 0.970292 \\
        7.0 & 21.654276 & 0.558010 & -34.785660 \\
        8.0 & 43.488829 & 85.452632 & -28.868817
\end{tabular}
\end{ruledtabular}
\end{table}

\begin{table}[h!]
\caption{\label{tab:table2} Ground state energy (in Ha) of the He-He system}
\begin{ruledtabular}
\begin{tabular}{c|ccc}
    \diagbox[height=1.8\line]{$r$ (a.u.)}{$\tau$ (\textcolor{teal}{unit})}& 0.0 & 1.0 & 2.0\\
    \hline
    1.0 & 1.000000 & 1.000000 & 5.739538 \\
    2.0 & 1.266281 & 1.134311 & -0.966946 \\
    3.0 & 0.880808 & 0.147328 & -0.206244 \\
    4.0 & 3.755382 & 0.136923 & 1.000000 \\
    5.0 & 1.000000 & 0.999847 & 0.415453 \\
    6.0 & 6.604936 & 1.000000 & -1.709232 \\
    7.0 & 1.000008 & 1.000000 & 9.223832 \\
    8.0& 198.102785 & 1.000000 & 20.246740
\end{tabular}
\end{ruledtabular}
\end{table}

>>>>

\title{Energy spectrum of diatomic molecules\\ in the Quantum Drude Oscillator model}
%\thanks{}

@@

        Let us move to the center of mass frame, in which the dynamics reduces to that of an effective reduced particle:
        \begin{subequations}
        \begin{align}
            &\vec x:=\vec x_1 - \vec x_2\\
            &m_1\vec x_1 + m_2\vec x_2 = \vec 0\,.
        \end{align}
        \end{subequations}
        In terms of the reduced variable, a straightforward computation gives us:
        \begin{subequations}
        \begin{align}
            V_0(\vec x)=\,&-\frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{ x^2-3(\vec x\cdot\hat r)^2}{4\pi\epsilon_0r^3}\,,\\
            V_1(\vec x)=\,&\frac{3m_1m_2}{2(m_1+m_2)^2}\cdot\frac{(\vec x\cdot\hat r)\left(3x^2 - 5(\vec x\cdot\hat r)^2\right)}{4\pi\epsilon_0r^4}\,,\\
            V_2(\vec x)=\,&\frac{m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{4(m_1+m_2)^4}\times\\
            &\times\frac{3x^4 - 30(\vec x\cdot\hat r)^2 x^2+35(\vec x\cdot\hat r)^4}{4\pi\epsilon_0r^5}\nonumber\,.
        \end{align}
        \end{subequations}
        Let us now assume that the system is effectively one-dimensional, aligned along the direction on which the centers of the two QDOs are sitting. One obtain the following one-dimensional potentials for an effective particle of mass $m_1m_2 / (m_1+m_2)$:
        \begin{subequations}
        \begin{align}
            V_0(x)&=\frac{2m_1m_2}{(m_1+m_2)^2}\cdot\frac{2 x^2}{4\pi\epsilon_0r^3}\,,\\
            %V_1(x)&=-\frac{3m_1m_2}{(m_1+m_2)^2}\cdot\frac{\text{sgn}(\vec x\cdot\hat r)|x|x^2}{4\pi\epsilon_0r^4}=\mathbin{\textcolor{teal}{-\frac{3m_1m_2}{(m_1+m_2)^2}\cdot\frac{x^3}{4\pi\epsilon_0r^4}}}\,,\\
            V_1(x)&=-\frac{3m_1m_2}{(m_1+m_2)^2}\cdot\frac{x^3}{4\pi\epsilon_0r^4}\,,\\
            V_2(x)&=\frac{2m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{(m_1+m_2)^4}\cdot\frac{x^4}{4\pi\epsilon_0r^5}\,.
        \end{align}
        \end{subequations}
        Including the contribution of the harmonic potentials present in (\ref{eq:full_QDO_Hamiltonian}) and coupling the system to an external electric field, one finally obtain the following total one-dimensional potential:
        \begin{equation}
        \label{eq:quartic_potential}
            V(x) = \alpha x + \beta x^2 + \gamma x^3 + \delta x^4 \,,
        \end{equation}
        with the following expression of the coupling constants:
        \begin{subequations}
        \label{eq:coupling_constants}
        \begin{align}
            \alpha &= -\frac{m_1 q_2 + m_2 q_1}{m_1+m_2}\cdot \mathcal E\\
            \beta &= \frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{2q_1 q_2}{4\pi\epsilon_0r^3} + \frac{1}{2}\cdot\frac{m_1m_2}{m_1+m_2}\cdot\frac{m_1\omega_2^2+m_2\omega_1^2}{m_1+m_2}\\
            \gamma&=-\frac{m_1m_2}{(m_1+m_2)^2}\cdot\frac{3 q_1 q_2}{4\pi\epsilon_0r^4}\,,\\
            \delta&=\frac{2m_1m_2(2m_1^2+3m_1m_2+2m_2^2)}{(m_1+m_2)^4}\cdot\frac{q_1 q_2}{4\pi\epsilon_0r^5}\,.
        \end{align}
        \end{subequations}

@@

    In what follows, we define $k^2:=\frac{m_1+m_2}{m_1 m_2}\frac{\hbar^2}{2}$.
    Let us consider the time-independent Schr\"odinger equation
    \begin{equation}
        \left[\frac{\text{d}^2}{\text{d}x^2} + \frac{\epsilon - V(x)}{k^2}\right]\psi(x) = 0\,,
    \end{equation}
    where the potential $V$ is provided by (\ref{eq:quartic_potential}) and (\ref{eq:coupling_constants}).

    Following \cite{Lay_1998}, we solve the problem separately on the positive and negative real axis, and impose that the two solutions (positive and negative) $\psi_\pm$ and their first derivative match at the origin. For brievety reasons, we treat explicitely the case of the positive solution, the negative case being treated by perfoming the transformation $(x, \alpha, \beta, \gamma, \delta)\rightarrow (-x, -\alpha, \beta, -\gamma, \delta)$.

    Performing the following change of variable:
    \begin{equation}
        \psi_+(x)=\exp\left[\frac{\gamma^2-4\beta\delta}{8k\delta^{3/2}}\,x-\frac{\gamma}{4k\delta^{1/2}}\,x^2-\frac{\delta^{1/2}}{3k}\,x^3\right]\phi_+(x)
    \end{equation}
    we obtain:
    \begin{widetext}
    \begin{equation}
    \begin{split}
        &\frac{\text{d}^2\phi_+}{\text{d}x^2}(x)+\left[\frac{\gamma^2-4\beta\delta}{4k\delta^{3/2}}-\frac{\gamma}{k\delta^{1/2}}\,x-\frac{2\delta^{1/2}}{k}\,x^2\right]\frac{\text{d}\phi_+}{\text{d}x}(x)\\
        &+\left[\frac{16 \delta ^2 \left(\beta ^2+4 \delta  \epsilon \right)-8 \beta  \gamma ^2 \delta +\gamma ^4-32 \gamma  \delta ^{5/2} k}{64 \delta ^3 k^2}-\frac{8 \alpha  \delta ^2-4 \beta  \gamma  \delta +\gamma ^3+16 \delta ^{5/2} k}{8 \delta ^2 k^2}\,x\right]\phi_+(x)=0\,,
    \end{split}
    \end{equation}
    \end{widetext}
    which is known as the so-called \textit{tri-confluent Heun equation}.
    We further perform the following change of variable, the reason why to become clear soon:
    \begin{equation}
        \phi_+(x)=(1+x)^{\frac{4 \beta  \gamma  \delta -\gamma ^3-8 \delta ^2 \left(\alpha +2k \delta^{1/2} \right)}{16k \delta ^{5/2}}}f_+(x)\,,
    \end{equation}
    as well as map the positive real axis to the segment $[0,1]$ through the change of argument
    \begin{equation}
        z=\frac{x}{x+1}\,,
    \end{equation}
    and look for a solution of the obtained ODE in terms of a series expansion at zero of the form:
    \begin{equation}
        f_+(z)=\sum_{n\geq 0}c^+_nz^n\,.
    \end{equation}
    Plugging the ansatz into the differential equation, we obtain the following recurrence relation between the coefficients:
    \begin{widetext}
    \begin{equation}
    \begin{split}
        &(n+1)(n+2) c^+_{n+2}+(n+1)(A -4n)c^+_{n+1} +[6(n-1) n+B n+E]c^+_n \\
        &+[C (n-1)+F-4(n-2) (n-1)]c^+_{n-1} + [(n-3) (n-2)+D (n-2)+G]c^+_{n-2}=0\ \ \ (n\geq 2)
    \end{split}
    \end{equation}
    \end{widetext}
    where the various coefficients are listed in full details in appendix. The coefficients $c^+_0$ and $c^+_1$ are to be fixed through boundary conditions, and the coeffients $c^+_2$ and $c^+_3$ are in turn fixed by the following relations
    \begin{equation}
    \begin{split}
        c^+_2&=-Ec^+_0-Ac^+_1\,,\\
        c^+_3&=-\frac{ F c^+_0+(B+E)c^+_1 +2 (A-4)c^+_2 }{6}\,.
    \end{split}
    \end{equation}
    Let us similarly denote by $c^-_n$ the coefficients in the series expansion corresponding to the solution on the negative real axis. It has been shown in \cite{Lay_1998} and \cite{doi:10.1063/1.531962} that the smooth matching of the two solutions at the origin can be achieved by imposing the so-called \textit{dominant solution} to the above defined difference equations vanishes. The conditions
    \begin{equation}
        \psi_+(0) = \psi_-(0)\ \ \ \text{and}\ \ \ \frac{\text{d}\psi_+}{\text{d}x}(0) = \frac{\text{d}\psi_-}{\text{d}x}(0)
    \end{equation}
    lead to constraints on the initial conditions for the difference equations, namely that $c^+_0 = c^-_0$ and:
    \begin{equation}
    \label{eq:initial_constraint}
    \begin{split}
        &c^+_1-c^+_0 \left[1+\frac{8 \delta ^2 (\alpha +\beta )-2 \gamma  \delta  (\gamma+2 \beta )+\gamma ^3}{16 \delta ^{5/2} k}\right]\\
        & = -c^-_1+c^-_0 \left[1-\frac{8 \delta ^2 (\alpha -\beta )+2 \gamma  \delta  (\gamma -2 \beta )+\gamma ^3}{16 \delta ^{5/2} k}\right]
    \end{split}
    \end{equation}
    The two sides of relation (\ref{eq:initial_constraint}) are therefore equal to some quantity $s$, which parameterizes the solution. Now, the vanishing of the dominant solution can be achieved by imposing that the series coefficients $c^+_n(\epsilon, s)$ and $c^-_n(\epsilon, s)$ are identically $0$ asymptotically in $n$. This provides a pair of algebraic equations in the energy variable $\epsilon$ and the \textit{shift} variable $s$.

    Let us make a few remarks at this stage. Imposing the vanishing of $c^\pm_n$ asymptotically in $n$ means that practically one should fix a large value of $n$, denote it $N$, and impose that the polynomials $c^\pm_N$ vanish. One observes that $c^-_N$ is a linear function of $s$. One can therefore solve explicitely for the shift variable in terms of the energy $\epsilon$. Plugging back into the other constraint, one obtains a rational function of the form:
    \begin{equation}
        c^+_N(\epsilon, s(\epsilon)) = \frac{P_N(\epsilon)}{Q_N(\epsilon)}
    \end{equation}
    where $P_n$ and $Q_N$ are polynomial functions of $\epsilon$, whose degree $d_N$ is an non-decreasing function of $N$. The $N$ zeros $\left\{\epsilon_i\right\}_{i=1}^N$ correspond then to first $N$ energy levels of the system, provided they are not complex and that $Q_N$ does not admit a zero of the same order.

@@

\begin{table}[h!]
\caption{\label{tab:table1} Ground state energy (in Ha) of the H-H system}
\begin{ruledtabular}
\begin{tabular}{c|ccc}
    \diagbox[height=1.8\line]{$r$ (a.u.)}{$\mathcal E$ (a.u.)}& 0.0 & 1.0 & 2.0\\
    \hline
    1.0 & 5.428021 & 1.000000 & 3.673658 \\
    2.0 & 0.531071 & 0.210137 & 0.872826 \\
    3.0 & 1.532129 & 0.957917 & 0.999836 \\
    4.0 & 1.968642 & 0.999963 & 1.000023 \\
    5.0 & 0.031075 & 1.000000 & 1.014180 \\
    6.0 & 0.978062 & 5.375749 & 0.970292 \\
    7.0 & 21.654276 & 0.558010 & -34.785660 \\
    8.0 & 43.488829 & 85.452632 & -28.868817
\end{tabular}
\end{ruledtabular}
\end{table}

\begin{table}[h!]
\caption{\label{tab:table2} Ground state energy (in Ha) of the He-He system}
\begin{ruledtabular}
\begin{tabular}{c|ccc}
    \diagbox[height=1.8\line]{$r$ (a.u.)}{$\mathcal E$ (a.u.)}& 0.0 & 1.0 & 2.0\\
    \hline
    1.0 & 1.000000 & 1.000000 & 5.739538 \\
    2.0 & 1.266281 & 1.134311 & -0.966946 \\
    3.0 & 0.880808 & 0.147328 & -0.206244 \\
    4.0 & 3.755382 & 0.136923 & 1.000000 \\
    5.0 & 1.000000 & 0.999847 & 0.415453 \\
    6.0 & 6.604936 & 1.000000 & -1.709232 \\
    7.0 & 1.000008 & 1.000000 & 9.223832 \\
    8.0& 198.102785 & 1.000000 & 20.246740
\end{tabular}
\end{ruledtabular}
\end{table}

    \begin{figure}[h!]
        \centering
            \includegraphics[scale=0.33]{plus_minus_coefs}
            \caption{\label{fig:coefs} Intersection loci of the three surfaces $z=c^+_{20}(\epsilon, s)$ and $z=c^-_{20}(\epsilon, s)$ and $z=0$.}
    \end{figure}
